---
title: "purrr::A_Primer"
author: "Karl Hailperin"
date: "November 30, 2017 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---
# What is purrr?

- Package created by Lionel Henry and co-authored by Hadley Wickham, both of RStudio
- Part of the broader tidyverse
- Contains a variety of functions to make programing easier, mostly by reducing for loops


---
# So What?

Which would you rather write?

This:
```{r ugly_mean}
x <- 1:5
sum <- 0
for (i in seq(length(x))){
        sum <- sum + x[i]
}
sum/length(x)
```

Or this:
```{r clean_mean}
x <- 1:5
mean(x)
```


---
# The map() Family, Part 1: The basics
- Basic form: map(.x, .f, ...)
- Creates results like this:

```
results = vector("list", length(.x))
for (i in seq(length(.x))){
        results[i] <- .f(.x[i], ...)
}
results
```

```{r map_example_1}
library(purrr)
numbers <- 1:3
map(numbers, sqrt)
```


---
# The map() Family, Part 2: map_*

- map() returns a list, but the map_* family returns a vector
- For example, map_chr returns a character vector
- Other suffixes include lgl (booleans), int (integers), dbl (numeric)


```{r map_example_2}
map_dbl(numbers, sqrt)
```


---
# The map() Family, Part 3: Other ways to use .f

- If .x is a nested list, using a number for .f returns the nth element of each element of .x
- If the nested lists are named, you can string inputs to get that input


---
# The map() Family, Part 3: Other ways to use .f
```{r map_example_3}
people <- list(c(first_name = "Kristen", last_name = "Bell"), 
               c(first_name = "Ted", last_name = "Danson"), 
               c(first_name = "William", middle_name = "Jackson", last_name = "Harper"))

map_chr(people, 2)
map_chr(people, "last_name")
map_chr(people, 3, .default = NA)
```

---
# The map() Family, Part 4: anonymous functions

- You can add more complex functions in .f with ~ notation

```{r map_example_4, eval = FALSE}
plots <- list(Plot_1 = plot_1, Plot_2 = plot_2, Plot_3 = plot_3)
plot_paths <- map_chr(names(plots), ~ paste0("Plots/", ., ".png"))
```



---
# The map() Family, Part 5: map2()

- map2 is like map, but takes two lists, and has the form map2(.x, .y, .f, ...)
- has map2_* functions
```{r map_example_5}
vector_1 <- c("Spite", "Research", "Boom", "Far")
vector_2 <- c("Malice", "Development", "Bust", "Away")
map2_chr(vector_1, vector_2, paste, sep = " and ")
```


---
# The map() Family, Part 6: pmap()

- Instead of creating abitrarily many map[number]() functions, purrr has pwalk(.l, .f, ...)
- .l is a list of lists
- There are pmap_[type] functions
```{r map_example_6}
observations <- 1:3
sizes <- 6 # will be recycled
probabilities <- c(0.6, 0.7, 0.8)
pmap(list(observations,sizes, probabilities), rbinom)
```


---
# The map() Family, Part 7: walk()

- walk() is map, but for functions that like write_csv() where we only care about the side effects (the file creation, in this case)

```{r map_example_8, eval = FALSE}
walk2(plot_paths, plots, ggsave)
```

---
# The map() Family, Part 8a: modify_depth()

- modify_depth(.x, .depth, .f, ...) is for mapping over elements buried in a nested list
- .depth = 0 is the list itself
- .depth = 1 is the elements of the list (modify_depth(.x, 1, .f is the same as map(.x, ,f)))
- .depth = 2 is the elements of those lists (assuming .x is a list of  lists)

```{r map_example_11a}
casts <- list(
        the_good_place = people,
        crazy_ex_girlriend = list(c(first_name = "Rachel", last_name = "Bloom"), 
                                       c(first_name = "Vincent", last_name = "Rodriguez"), 
                                       c(first_name = "Donna", middle_name = "Lynne", last_name = "Champlin"))
)
```


---
#The map() Family, Part 8b: modify_depth()

```{r map_example_11b}
map(casts, toupper)
modify_depth(casts, 2, toupper)
```


---
# reduce()

- reduce(.x, .f) uses .f to reduce .x to one element, starting with the first 2, than the output and the third,etc.
- If x is a 3-element list/vector, `map(x,f) = f(f(x1,x2),x3)`
- Example: `x = c(1,2,3)`, ``map(x, `+`) = (1 +2) + 3``
- Like map, reduce also has a reduce_2 version
- reduce_right() works like reduce, but from the right
- accumulate(): like reduce(), but returns a vector with all the intermediate values in addition to the final one
    + Example: `x = c(1,2,3)`, ``accumulate(x, `+`)`` returns `r c(1,3,6)`
    + While there's an accumulate_right(), there's no accumulate2()
    
```{r reduce_example}
one_to_hundred <- 1:100
reduce(one_to_hundred, `+`)
```


---
# safely()

- safely(.f) returns a version of .f that handles errors well
```{r safely_example}
safe_sqrt <- safely(sqrt)
safe_sqrt(4)
safe_sqrt("a")
```

- Don't like NULL? Use the otherwise argument.
- .f can use ~ notation.

---
# Use cases
- list columns
- important for working with `tidyr::nest`
- Relatedly, working with cross-validation sets created by [`{rsample}`](https://rsample.tidymodels.org/)

---
# Other Resources Because You Found This Talk Unhelpful
- Charlotte Wickham: [https://github.com/cwickham/purrr-tutorial/blob/master/slides.pdf](https://github.com/cwickham/purrr-tutorial/blob/master/slides.pdf)
- Julia Silge:[https://jennybc.github.io/purrr-tutorial/index.html](https://jennybc.github.io/purrr-tutorial/index.html)
    + [discussion of list columns](https://jennybc.github.io/purrr-tutorial/ls13_list-columns.html)

